---
title: "MurineMS"
output: html_document
---

#First we load some libraries

```{r}
library(phyloseq)
library(phytools)
library(ggplot2)
library(dendextend)
library(zCompositions)
library(compositions)
library(vegan)
library(ALDEx2)
```
#Then we separate otu table and taxonomy table that came together as a result of 
#Qiime processing; otu_table should be included in the repository, as well as sample metadata and tree
```{r}
otu_table_HIT <- read.csv("data/otu_table_HIT.csv")
tax_raw_HIT <- otu_table_HIT[,12]
rownames(otu_table_HIT)<- otu_table_HIT[,1]
###
tax_rat_HIT <- data.frame(OTU=rownames(otu_table_HIT), p=NA, c=NA, o=NA, f=NA, g=NA, s=NA)
for (i in 1:length(tax_raw_HIT)){
  tax_rat_HIT$p[i]<- unlist(strsplit(as.character(tax_raw_HIT[i]), ";"))[1]
  tax_rat_HIT$c[i]<- unlist(strsplit(as.character(tax_raw_HIT[i]), ";"))[2]
  tax_rat_HIT$o[i]<- unlist(strsplit(as.character(tax_raw_HIT[i]), ";"))[3]
  tax_rat_HIT$f[i]<- unlist(strsplit(as.character(tax_raw_HIT[i]), ";"))[4]
  tax_rat_HIT$g[i]<- unlist(strsplit(as.character(tax_raw_HIT[i]), ";"))[5]
  tax_rat_HIT$s[i]<- unlist(strsplit(as.character(tax_raw_HIT[i]), ";"))[6]
}
row.names(tax_rat_HIT)<- tax_rat_HIT$OTU
tax_rat_HIT <- tax_rat_HIT[,-1]
tax_rat_HIT <- tax_table(tax_rat_HIT)
#
row.names(tax_rat_HIT) <- rownames(otu_table_HIT)
#
otu_rat_HIT <- otu_table(otu_table_HIT[,-c(1,12)], taxa_are_rows=TRUE)
head(otu_rat_HIT)
```
#Next we import tree from HIT.db, and setup sample data, combining all into the phyloseq object
#here we should include path to the tree
```{r}
phy<-read.newick("data/HITdb.tre")
phy<-collapse.singles(phy)
phy<-multi2di(phy)
treRatHIT <-phy
#tree imported
if (min(phy$edge.length) == 0) { phy$edge.length = phy$edge.length + 0.00001 }
###### now sample data
qiimeMap <- read.delim("data/qiimeMap.txt")
head(qiimeMap)
qiimeRat <- qiimeMap[qiimeMap$Description=="rat",]
rownames(qiimeRat) <- qiimeRat$SampleID
sampleRat <- sample_data(qiimeRat)
#######
phyloRatHIT <- phyloseq(otu_rat_HIT, sampleRat, tax_rat_HIT, treRatHIT)
summary(phyloRatHIT)
class(phyloRatHIT)
phyloRatHIT

```

## Now we are putting more metadata
```{r}
sample_data(phyloRatHIT)$AntibioticTreatment <- c("antibiotic-free","antibiotic","antibiotic","antibiotic-free","antibiotic-free","antibiotic-free","antibiotic-free","antibiotic","antibiotic","antibiotic")
sample_data(phyloRatHIT)$ChildTime <- c("-28.dpi","12.dpi","dams","31-35.dpi","dams", "12.dpi","0.dpi","-28dpi","0.dpi","31-35.dpi")
sample_data(phyloRatHIT)$ChildTime <- factor(sample_data(phyloRatHIT)$ChildTime, levels=c("dams", "-28.dpi", "0.dpi", "12.dpi", "31-35.dpi"))
```
## rarefying and plotting alpha diversity measures
```{r}
set.seed(123)
alpha_meas = c("Shannon")
phyloRatHITrar <- rarefy_even_depth(phyloRatHIT)
sample_data(phyloRatHITrar)$time_point<-
ifelse(sample_data(phyloRatHITrar)$ChildTime=="razdvoj", "-28 dpi", 
                                  ifelse(sample_data(phyloRatHITrar)$ChildTime=="12dpi", "12 dpi",
                                                        ifelse(sample_data(phyloRatHITrar)$ChildTime=="mother", "dams",
                                                               ifelse(sample_data(phyloRatHITrar)$ChildTime=="3135dpi", "31-35 dpi",
                                                                      ifelse(sample_data(phyloRatHITrar)$ChildTime=="3135dpi", "31-35 dpi",
                                                                             ifelse(sample_data(phyloRatHITrar)$ChildTime=="0dpi", "0 dpi", NA))))))
p <- plot_richness(phyloRatHITrar,"time_point","AntibioticTreatment", measures=alpha_meas)
#some level ordering
sample_data(phyloRatHITrar)$time_point <- factor(sample_data(phyloRatHITrar)$time_point, levels=c("dams", "-28 dpi","0 dpi","12 dpi", "31-35 dpi"))

ggplot(p$data, aes(x=AntibioticTreatment, y=value, color=samples)) +geom_point(size=3)+geom_boxplot(data=p$data, aes(x=AntibioticTreatment, y=value, color=NULL), alpha=0.1)+theme_bw()+labs(x="", y="Shannon index")+theme(axis.text.x=element_text(size=18), axis.title.y=element_text(size=18))
```
#Next we do some ordination (unconstrained)
```{r}
uniRatHIT <- distance(phyloRatHITrar, "wunifrac")
#
PCoA.ratHIT <- ordinate(phyloRatHITrar, method="PCoA", distance=uniRatHIT)
plot_scree(PCoA.ratHIT, "Scree plot for Rat data, UniFrac/PCoA")
p12 <-plot_ordination(phyloRatHITrar, PCoA.ratHIT, axes=c(1, 2),"SampleID", color="AntibioticTreatment")+geom_point(size=5) + geom_path() + scale_colour_hue(guide = FALSE) 
```
#Now we use the same weighted unifrac distance to do the clustering
```{r}
rat.hclustHIT<-hclust(d=uniRatHIT,method="ward.D")
rat.dendHIT <- as.dendrogram(rat.hclustHIT)
###some label manipulation
labels(rat.dendHIT)
data.clust <- data.frame(sampleID = c("M8","M10","M1","M5","M9","M7","M6","M2","M4","M3"))
data.clust$AntibioticTreatment <- sample_data(phyloRatHITrar)[as.character(data.clust$sampleID),"AntibioticTreatment"]
data.clust$AntibCol <- ifelse(data.clust$AntibioticTreatment=="antibiotic-free", "deepskyblue2", "red1")
data.clust$ChildTime <- sample_data(phyloRatHITrar)[as.character(data.clust$sampleID),"ChildTime"]
#labels(rat.dendHIT) <- data.clust$ChildTime
labels(rat.dendHIT) <- c("dams", "-28.dpi", "0.dpi", "  31-35.dpi","-28.dpi", "dams","31-35.dpi", "0.dpi","12.dpi", "12.dpi")
labels_colors(rat.dendHIT)<- data.clust$AntibCol
plot(rat.dendHIT)#this is the code for hierarchical clustering tree from the paper
```
#now exploratory phylogenetic tree
```{r}
RatHIT.imun <- subset_taxa(phyloRatHITrar, ta5 %in% c(" Peptococcus", " Peptoclostridium"," Parasutterella", " Collinsella", " Turicibacter"))
plot_tree(RatHIT.imun, color="ChildTime", size="Abundance",shape="AntibioticTreatment", nodelabf=nodeplotblank, ladderize="left", base.spacing=0.04,label.tips="ta6" )
#this is the piece of phylogenetic tree, from the paper
```
#Now we are entering the realm of compositional framework
```{r}
d.pro.HIT <- cmultRepl(t(otu_table(phyloRatHIT)), method="CZM", output="counts")
#this is the way to account for zero counts
d.pro.HIT <- t(d.pro.HIT)
d.pro.HIT <- as.data.frame(d.pro.HIT)
d.pro.HIT$genus <- tax_table(phyloRatHIT)[,"ta5"]
#now aggregate, and do the clr transformation
genusRatHIT <-aggregate(.~genus, data=d.pro.HIT[,1:11], sum)
d.clr.abund.HIT <- t(apply(genusRatHIT[,-1], 2, function(x){log(x) - mean(log(x))}))
colnames(d.clr.abund.HIT) <- genusRatHIT[,1]
```
#Now some constrained ordination, by antibiotic usage
```{r}
rda.part <- rda(d.clr.abund.HIT~AntibioticTreatment, data=data.frame(sample_data(phyloRatHIT)))
##
scor <- scores(rda.part, display=c("sp", "cn", "bp", "wa"), scaling=2)
##
numeric_centroids <- data.frame(scor$centroids)
numeric_centroids$AntibioticTreatment <- c("antibiotic", "antibiotic-free")
#ggplot(numeric_centroids, aes(x=RDA1, y=PC1))+geom_text(aes(label=AntibioticTreatment))
##
species_centroids <- data.frame(scor$species)
species_centroids$species_names <- rownames(species_centroids)
##introducing criteria to avoid clutter in the middle of the plot
species_centroids$kriterijum <- sqrt(species_centroids$RDA1^2+species_centroids$PC1^2)
#
index.spec <- species_centroids$kriterijum > 0.75
species_centroids.over <- species_centroids[index.spec,]
#
site_scores <- data.frame(scor$sites)
site_scores$ChildTime <- sample_data(phyloRatHIT)$ChildTime
site_scores$SampleID <- sample_data(phyloRatHIT)$SampleID
site_scores$AntibioticTreatment<- sample_data(phyloRatHIT)$AntibioticTreatment
#
df_ell.antib<- data.frame()
veganCovEllipse <- function(cov, center=c(0,0), scale=1, npoints=100){
  theta <- (0:npoints)*2*pi/npoints
  Circle <- cbind(cos(theta), sin(theta))
  t(center+scale*t(Circle%*%chol(cov)))
}
site_scores$AntibioticTreatment <- factor(site_scores$AntibioticTreatment)
for (g in levels(site_scores$AntibioticTreatment)){
  df_ell.antib<- rbind(df_ell.antib,cbind(as.data.frame(with(site_scores[site_scores$AntibioticTreatment==g,], veganCovEllipse(cov.wt(cbind(RDA1,PC1),wt=rep(1/length(RDA1), length(RDA1)))$cov, center=c(mean(RDA1), mean(PC1))))),AntibioticTreatment=g))}
#
RDA.plot.genus <-ggplot(site_scores, aes(x=RDA1, y=PC1))+geom_point(aes(colour=AntibioticTreatment, shape=AntibioticTreatment), size=5)+geom_path(data=df_ell.antib, aes(x=RDA1, y=PC1, group=AntibioticTreatment, colour=AntibioticTreatment), alpha=0.3, linetype=1)+
  geom_segment(data=species_centroids.over, aes(x=0, xend=2*RDA1, y=0, yend=8.5*PC1), arrow=arrow(length=unit(0.25, "cm")), colour="grey4", alpha=0.4, linetype=2)+
  geom_text(data=species_centroids.over, aes(x=2.5*RDA1, y=9.3*PC1,label=species_names), size=4, colour="black")+theme_bw()
RDA.plot.genus
####formal testing
anova(rda.part, by="terms")
```
#and now some testing with ALDEx
```{r}
d.aldex <- data.frame(otu_table(phyloRatHITrar))
d.aldex$genus <- tax_table(phyloRatHITrar)[,"ta5"]
index.NA.genus <- !is.na(d.aldex$genus)
#removing taxa not annotated on genus level
d.genus <- d.aldex[index.NA.genus,]
genusRatHIT <-aggregate(.~genus, data=d.genus[,1:11], sum)
rownames(genusRatHIT) <- genusRatHIT$genus
##
conds.aldex <- c("antibiotic-free","antibiotic","antibiotic-free","antibiotic","antibiotic-free","antibiotic-free","antibiotic-free","antibiotic","antibiotic","antibiotic")
x <- aldex.clr(genusRatHIT[,2:11], conds.aldex,mc.samples=128, verbose=TRUE)
#calculating p value for each replicate
x.t <- aldex.ttest(x, conds.aldex)
x.e <- aldex.effect(x, conds.aldex, verbose=TRUE)
x.all <- data.frame(x.e, x.t)
sig <- which(x.all$wi.eBH <=0.05)
```
#No genus was found significantly associated with antibiotic usage


